---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# macklinear

<!-- badges: start -->
<!-- badges: end -->

The goal of macklinear is to ...

## Installation

1. Download and install/update R.

2. Open R and run:

``` {r packages}
install.packages("macklinear")
library(macklinear)
``` 

Or you can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("AU-R-Programming/Final_Project_Group_5")
```

## Tutorial

This is a basic example using the base R data set, cars, which shows you how to solve a common problem.
<br><br><br>
The cars data set has two variables, **speed** and **distance**. We will use the `my_lm` function of the `macklinear` package to fit a regression and estimate the fit of our explanatory variable to our response variable.
<br><br><br>
```{r example}
library(macklinear)

head(cars)

plot(cars)
```
<br><br><br>
Our function, `my_lm()` needs four paramaters: 
`response`, `covariate`, `alpha` & `method`
<br><br><br>
We can see that the y variable, or the response variable, is distance. The variable **"dist"** is our `response` parameter in our function.
<br><br><br>
The x variable, or the explanatory variable, is speed. The variable **"speed"** is our `covariate` parameter in our function. 
<br><br><br>
`Alpha` is our level of significance set for our function. Alpha is generally set at 0.05, indicating a 95% confidence interval. If you do not set 0 < alpha < 1, you will have a warning generated as the true value of alpha must be between 0 and 1. Additionally, if you do not set 0.01 < alpha <0.1, you will have a warning generated. Alpha is typically between 0.01 and 0.1, therefore you should consider using a different alpha value.
<br><br><br>
The final paramater in the function, `method` is determining the method with which confidence intervals for β will be generated. You can choose to use a "boostrap" or "asymptotic" method for generating confidence intervals. 
<br><br><br>
We will now use the `my_lm()` function to analyze the relationship between **speed and distance**, with a **95%** confidence interval and a **bootstrap** approach for generating confidence interval. 

``` {r example 2}

my_lm(cars$dist, cars$speed, alpha = 0.05, method = "bootstrap")

```
<br><br><br>
Our function returns multiple statistics regarding the relationship between our x and y variables, **speed** and **distance**. 
<br><br><br>
`beta`

`sigma2` is an estimate of the residual variance. 

`variance_beta` is an estimate of the variance of the estimated beta.

`ci` is the confidence interval based on the alpha set in the function paramaters. 

`mspe` 

`ssm`

`sse`

`f.stat`

`residual`

`y.hat` 


